### 2023-09-09 draft.Rmd
updating simulations

```{=tex}
\newcommand{\diag}{\mathrm{diag}}
\newcommand{\tr}{\mathrm{Tr}}
\newcommand{\blockdiag}{\mathrm{blockdiag}}
\newcommand{\indep}{\stackrel{\mathrm{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\mathrm{iid}}{\sim}}
\newcommand{\Bernoulli}{\mathrm{Bernoulli}}
\newcommand{\Betadist}{\mathrm{Beta}}
\newcommand{\BG}{\mathrm{BernoulliGraph}}
\newcommand{\Uniform}{\mathrm{Uniform}}
\newcommand{\PABM}{\mathrm{PABM}}
\newcommand{\RDPG}{\mathrm{RDPG}}
\newcommand{\GRDPG}{\mathrm{GRDPG}}
\newcommand{\Multinomial}{\mathrm{Multinomial}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\as}{\stackrel{\mathrm{a.s.}}{\to}}
\newcommand{\ER}{\text{Erd\"{o}s-R\'{e}nyi}}
\newcommand{\SBM}{\mathrm{SBM}}
\newcommand{\DCBM}{\mathrm{DCBM}}
\newcommand{\rank}{\mathrm{rank}}
\newcommand{\MBM}{\mathrm{MBM}}
```

@@

@@

HCP data

```{r hcp-eda, cache = TRUE}
data.dir <- '~/Downloads/repeated_10_scale_125/'
graph.list <- lapply(dir(data.dir), function(x) {
  graph <- igraph::read.graph(file.path(data.dir, x), format = 'graphml')
  A <- graph %>% 
    igraph::as_adjacency_matrix() %>% 
    as.matrix()
  z <- graph %>% 
    igraph::get.vertex.attribute() %>% 
    {.$dn_hemisphere} %>% 
    as.factor()
  return(list(A = A, z = z, id = substr(x, 1, 6)))
})

i <- 1
graph <- graph.list[[i]]
A <- graph$A
z <- graph$z
Xhat <- embedding(A, 2, 0, scale = TRUE)
clustering <- manifold.clustering(Xhat, 
                                  degree = 1,
                                  initialization = as.numeric(z),
                                  intercept = FALSE)
plot.estimated.curves(Xhat, clustering)
```

```{r hcp-gender, cache = TRUE}
hcp.df <- readr::read_csv('~/Documents/postdoc-first-project/hcp-eda.csv')
hcp.df %>% 
  na.omit() %>% 
  GGally::ggpairs(columns = 1:4, aes(colour = gender))
```
```{r hcp-age, cache = TRUE}
hcp.df <- readr::read_csv('~/Documents/postdoc-first-project/hcp-eda.csv')
hcp.df %>% 
  na.omit() %>% 
  dplyr::filter(age != '36+') %>% 
  GGally::ggpairs(columns = 1:4, aes(colour = age))
```

>>>>

```{=tex}
\newcommand{\diag}{\mathrm{diag}}
\newcommand{\tr}{\mathrm{Tr}}
\newcommand{\blockdiag}{\mathrm{blockdiag}}
\newcommand{\indep}{\stackrel{\mathrm{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\mathrm{iid}}{\sim}}
\newcommand{\Bernoulli}{\mathrm{Bernoulli}}
\newcommand{\Betadist}{\mathrm{Beta}}
\newcommand{\BG}{\mathrm{BernoulliGraph}}
\newcommand{\Uniform}{\mathrm{Uniform}}
\newcommand{\PABM}{\mathrm{PABM}}
\newcommand{\RDPG}{\mathrm{RDPG}}
\newcommand{\GRDPG}{\mathrm{GRDPG}}
\newcommand{\Multinomial}{\mathrm{Multinomial}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\as}{\stackrel{\mathrm{a.s.}}{\to}}
\newcommand{\ER}{\text{Erd\"{o}s-R\'{e}nyi}}
\newcommand{\SBM}{\mathrm{SBM}}
\newcommand{\DCBM}{\mathrm{DCBM}}
\newcommand{\rank}{\mathrm{rank}}
\newcommand{\MBM}{\mathrm{MBM}}
\newcommand{\Poisson}{\mathrm{Poisson}}
```

@@

1. Draw angles $\theta_1, ..., \theta_N \iid \Uniform(\pi / 6, \pi / 3)$.
2. For each $k = 1, ..., N$, 
    i. Draw $t_1, ..., t_n \iid \Uniform(0, 1)$;
    ii. Draw $z_1, ..., z_n \iid \Multinomial(1/2, 1/2)$;
    iii. For each $i = 1, ..., n$, set $x_i = \begin{cases} \begin{bmatrix} t_i & 0 \end{bmatrix}^\top & z_i = 1 \\ \begin{bmatrix} t_i \cos \theta_k & t_i \sin \theta_k \end{bmatrix}^\top & z_i = 2 \end{cases}$;
    iv. Collect $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$ and draw $A \sim \RDPG(X)$;
    v. Set the response $y_k = \beta_0 + \beta_1 \theta_k$.

In this simulation, we set $\beta_0 = \beta_1 = 1$. 
The number of graphs was set to $N = 64$, and the number of vertices per graph was set to $n = 128, 256, 512, 1024$. 
For each $n$, we simulated 32 replicates.

```{r angle-reg-example, cache = TRUE}
n <- 512
a.vec <- c(1, 1)
b.vec <- c(1, 1)
theta.min <- pi / 6
theta.max <- pi / 3
beta1 <- 1
beta0 <- 1

z <- sample(seq(2), n, replace = TRUE)
t. <- rbeta(n, a.vec[z], b.vec[z])
theta <- runif(1, theta.min, theta.max)
y <- beta0 + beta1 * theta
x1 <- ifelse(z == 1, t., cos(theta) * t.)
x2 <- ifelse(z == 1, 0, sin(theta) * t.)
X <- cbind(x1, x2)
P <- X %*% t(X)
A <- draw.graph(P)
Xhat <- embedding(A, 2, 0)
clustering <- manifold.clustering(Xhat, 
                                  degree = 1,
                                  initialization = z,
                                  intercept = FALSE)
plot.estimated.curves(Xhat, clustering) + 
  scale_colour_brewer(palette = 'Set1')
```

```{r angle-reg-results, cache = TRUE}
out.df <- readr::read_csv('~/dev/multilayer-rdpg/simulations/angle-regression-sim.csv')
n.vec <- c(128, 256, 512, 1024)
N <- 64

ggplot(out.df) + 
  geom_boxplot(aes(x = n, y = mse * N, group = n)) + 
  scale_x_log10(breaks = sort(n.vec)) +
  scale_y_log10() +
  labs(x = 'number of vertices', y = 'SSE') + 
  theme_bw()
```

@@

```{r load-hcp}
load('~/Downloads/HCP/HCP-Y/DTI_fMRI/Data.RData')
ids <- dimnames(dti.countV2)[[3]]
```

In the first example, we analyzed fiber count data between brain regions from the Human Connectome Project (HCP). 
When analyzing these data as graphs, we denote the regions as vertices and the fiber counts between pairs of regions as weighted edges. 
A plausible statistical model for these data is to assume that the edge weights between pairs of vertices is Poisson distributed, i.e., the adjacency matrix is sampled as $A_{ij} \indep \Poisson(\Theta_{ij})$, where $\Theta \in \mathbb{R}_+^{n \times n}$ is a symmetric matrix of Poisson parameters.

In this dataset, there are $N = 136$ graphs (corresponding to individual subjects), each with $n = 68$ vertices (corresponding to brain regions). 
Analyzing these graphs as RDPGs reveals that the DCBM is a good candidate for these data (fig \ref{fig:hcp-ase}).

```{r hcp-ase, cache = TRUE, fig.cap = 'ASE of one graph from the HCP dataset. The lines are fitted via $K$-curves clustering using degree = 1. The outputted clusters correspond exactly to the left (red) and right (blue) hemispheres.', fig.width = 10, fig.height = 10}
id <- ids[1]
A <- dti.countV2[, , id] %>% 
  unique() %>% 
  t() %>% 
  unique()
z <- colnames(A) %>% 
  stringi::stri_sub(-1, -1) %>% 
  as.factor()

Xhat <- embedding(A, 2, 0, scale = TRUE)
clustering <- manifold.clustering(Xhat, 
                                  degree = 1,
                                  initialization = as.numeric(z),
                                  normalize = TRUE,
                                  intercept = FALSE,
                                  parallel = FALSE)
plot.estimated.curves(Xhat, clustering) + 
  scale_colour_brewer(palette = 'Set1') + 
  ggrepel::geom_label_repel(aes(x = Xhat[, 1],
                                y = Xhat[, 2],
                                label = colnames(A)))
```

For each graph, the

```{r hcp-gender, cache = TRUE}
hcp.df <- readr::read_csv('~/Documents/postdoc-first-project/hcp-eda.csv')
hcp.df %>% 
  na.omit() %>% 
  GGally::ggpairs(columns = 2:8, aes(colour = gender)) + 
  scale_colour_brewer(palette = 'Set1')
```
```{r hcp-age, cache = TRUE}
hcp.df <- readr::read_csv('~/Documents/postdoc-first-project/hcp-eda.csv')
hcp.df %>% 
  na.omit() %>% 
  dplyr::filter(age != '36+') %>% 
  GGally::ggpairs(columns = 2:8, aes(colour = age)) + 
  scale_colour_brewer(palette = 'Set1')
```

### 2023-09-09 draft.tex
updating simulations

\newcommand{\diag}{\mathrm{diag}}
\newcommand{\tr}{\mathrm{Tr}}
\newcommand{\blockdiag}{\mathrm{blockdiag}}
\newcommand{\indep}{\stackrel{\mathrm{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\mathrm{iid}}{\sim}}
\newcommand{\Bernoulli}{\mathrm{Bernoulli}}
\newcommand{\Betadist}{\mathrm{Beta}}
\newcommand{\BG}{\mathrm{BernoulliGraph}}
\newcommand{\Uniform}{\mathrm{Uniform}}
\newcommand{\PABM}{\mathrm{PABM}}
\newcommand{\RDPG}{\mathrm{RDPG}}
\newcommand{\GRDPG}{\mathrm{GRDPG}}
\newcommand{\Multinomial}{\mathrm{Multinomial}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\as}{\stackrel{\mathrm{a.s.}}{\to}}
\newcommand{\ER}{\text{Erd\"{o}s-R\'{e}nyi}}
\newcommand{\SBM}{\mathrm{SBM}}
\newcommand{\DCBM}{\mathrm{DCBM}}
\newcommand{\rank}{\mathrm{rank}}
\newcommand{\MBM}{\mathrm{MBM}}

@@

@@

HCP data

\begin{center}\includegraphics{draft_files/figure-latex/hcp-eda-1} \end{center}

>>>>

\newcommand{\diag}{\mathrm{diag}}
\newcommand{\tr}{\mathrm{Tr}}
\newcommand{\blockdiag}{\mathrm{blockdiag}}
\newcommand{\indep}{\stackrel{\mathrm{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\mathrm{iid}}{\sim}}
\newcommand{\Bernoulli}{\mathrm{Bernoulli}}
\newcommand{\Betadist}{\mathrm{Beta}}
\newcommand{\BG}{\mathrm{BernoulliGraph}}
\newcommand{\Uniform}{\mathrm{Uniform}}
\newcommand{\PABM}{\mathrm{PABM}}
\newcommand{\RDPG}{\mathrm{RDPG}}
\newcommand{\GRDPG}{\mathrm{GRDPG}}
\newcommand{\Multinomial}{\mathrm{Multinomial}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\as}{\stackrel{\mathrm{a.s.}}{\to}}
\newcommand{\ER}{\text{Erd\"{o}s-R\'{e}nyi}}
\newcommand{\SBM}{\mathrm{SBM}}
\newcommand{\DCBM}{\mathrm{DCBM}}
\newcommand{\rank}{\mathrm{rank}}
\newcommand{\MBM}{\mathrm{MBM}}
\newcommand{\Poisson}{\mathrm{Poisson}}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Draw angles
  \(\theta_1, ..., \theta_N \iid \Uniform(\pi / 6, \pi / 3)\).
\item
  For each \(k = 1, ..., N\),

  \begin{enumerate}
  \def\labelenumii{\roman{enumii}.}
  \tightlist
  \item
    Draw \(t_1, ..., t_n \iid \Uniform(0, 1)\);
  \item
    Draw \(z_1, ..., z_n \iid \Multinomial(1/2, 1/2)\);
  \item
    For each \(i = 1, ..., n\), set
    \(x_i = \begin{cases} \begin{bmatrix} t_i & 0 \end{bmatrix}^\top & z_i = 1 \\ \begin{bmatrix} t_i \cos \theta_k & t_i \sin \theta_k \end{bmatrix}^\top & z_i = 2 \end{cases}\);
  \item
    Collect
    \(X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top\) and
    draw \(A \sim \RDPG(X)\);
  \item
    Set the response \(y_k = \beta_0 + \beta_1 \theta_k\).
  \end{enumerate}
\end{enumerate}

In this simulation, we set \(\beta_0 = \beta_1 = 1\). The number of
graphs was set to \(N = 64\), and the number of vertices per graph was
set to \(n = 128, 256, 512, 1024\). For each \(n\), we simulated 32
replicates.

\begin{center}\includegraphics{draft_files/figure-latex/angle-reg-example-1} \end{center}

\begin{center}\includegraphics{draft_files/figure-latex/angle-reg-results-1} \end{center}

@@

In the first example, we analyzed fiber count data between brain regions
from the Human Connectome Project (HCP). When analyzing these data as
graphs, we denote the regions as vertices and the fiber counts between
pairs of regions as weighted edges. A plausible statistical model for
these data is to assume that the edge weights between pairs of vertices
is Poisson distributed, i.e., the adjacency matrix is sampled as
\(A_{ij} \indep \Poisson(\Theta_{ij})\), where
\(\Theta \in \mathbb{R}_+^{n \times n}\) is a symmetric matrix of
Poisson parameters.

In this dataset, there are \(N = 136\) graphs (corresponding to
individual subjects), each with \(n = 68\) vertices (corresponding to
brain regions). Analyzing these graphs as RDPGs reveals that the DCBM is
a good candidate for these data (fig \ref{fig:hcp-ase}).

\begin{figure}[H]

{\centering \includegraphics{draft_files/figure-latex/hcp-ase-1}

}

\caption{ASE of one graph from the HCP dataset. The lines are fitted via $K$-curves clustering using degree = 1. The outputted clusters correspond exactly to the left (red) and right (blue) hemispheres.}\label{fig:hcp-ase}
\end{figure}

For each graph, the
